# 概率论和数理统计 {#probability-statistics}

http://staff.ustc.edu.cn/~zwp/teach/Prob-Stat/probstat.htm

```{r}
# install.packages("TeachingDemos")
library("TeachingDemos")
```

## 掷骰子 {#roll-the-dice}

```{r}
dice(10, ndice=2, plot.it=T)
```


### Efron's dice

see: <http://mathworld.wolfram.com/EfronsDice.html>。

```{r}
 ed <- list( rep( c(4,0), c(4,2) ),
 rep(3,6), rep( c(6,2), c(2,4) ),
 rep( c(5,1), c(3,3) ) )
     
 tmp <- dice( 10000, ndice=4 )
 dice(10,ndice=4,plot.it=T)
 ed.out <- sapply(1:4, function(i) ed[[i]][ tmp[[i]] ] )
     
 mean(ed.out[,1] > ed.out[,2])
 mean(ed.out[,2] > ed.out[,3])
 mean(ed.out[,3] > ed.out[,4])
 mean(ed.out[,4] > ed.out[,1])
```



## Buffon's needle

```{r}
# install.packages("animation")
library(animation)
```


```{r}
oopt = ani.options(nmax = 1000, interval = 0)
opar = par(mar = c(3, 2.5, 0.5, 0.2), pch = 20, mgp = c(1.5, 0.5, 0))
buffon.needle()
```







## sample() - sampling from a finite set

```{r}
# sample(x, size, replace = FALSE, prob = NULL)

# Lotto genrator (sampling from 1:39 without replacement)

sample(39, 7) # or: sample(1:39, 7)

# Ten rows of LOTTO (surely one of these must win?!?)

replicate(10, sample(39, 7))
```

```{r}
# Use the argument prob, if you need to specify different
# probabilities for the different outcomes.

# Sometimes we want sampling with replacement; this is the
# same as drawing an i.i.d. sequence of values from the
# corresponding discrete distribution.

# Simulate dice throws

fair.die <- sample(1:6, 100, replace = TRUE)
table(fair.die)
```

Simulate dice throws with a loaded die.


```{r}
loaded.die <- sample(1:6, 100, replace = TRUE, prob = c(2, 2, 2, 2, 2, 5))
table(loaded.die)
```

## Discrete random variables and their distributions


1. Binomial 

```{r}
dbinom(x=2,size=20,prob=0.5)
pbinom(q=2,size=20,prob=0.5)
qbinom(p=0.4,size=20,prob=0.5)
rbinom(n=5,size=20,prob=0.5)


plot(dbinom(0:20,size=20,prob=0.5),type="h")
plot(dbinom(0:20,size=20,prob=0.8),type="h")
```

2. The Hypergeometric Distribution

```{r}
dhyper(x=2, m=10, n=30, k=6)
phyper(q=2, m=10, n=30, k=6)
qhyper(0.3, m=10, n=30, k=6)
rhyper(nn=10, m=10, n=30, k=6)
```

3. The Geometric Distribution,let X count the number of failures before the first successes

```{r}
dgeom(4,prob=0.8)
pgeom(4, prob = 0.8)
qgeom(0.4,prob=0.8)
rgeom(10,prob=0.8)


plot(dgeom(0:20,prob=0.5),type="h")
plot(dgeom(0:20,prob=0.8),type="h")
```

4. The Negative Binomial Distribution, let X count the number of failures before r successes

```{r}
dnbinom(x=5,size=3,prob=0.4)   
pnbinom(5,size=3,prob=0.4)
qnbinom(0.5,size=3,prob=0.4)
rnbinom(n=10,size=3,prob=0.4)

plot(dnbinom(0:20,size=5,p=0.5),type="h")
```

5. Poisson distributino

```{r}
dpois(x=0,lambda=2.4)
ppois(q=10,lambda=2.4)
qpois(p=0.9,lambda=2.4)
rpois(n=10,lambda=2.4)

plot(dpois(0:20,lambda=1),type="h")
```


```{r}
x <- 0:20
plot(x, ppois(x, 1), type="s", lty=1,ylab="F(x)", main="Poisson approx of binomial")
lines(x, pbinom(x, 100, 0.01),type="s",col=2,lty=2)
legend("bottomright",legend=c("Poisson","Binomial"),lty=1:2,col=1:2)
```

6.  Poisson and normal approximation of binomial probabilities, with estimated parameters

```{r}

#P(X<=k)=pbinom(k,n,p)
#Poisson approximation: P(X<=k) app ppois(k,np)
#Normal approximation: P(X<=k) app pnorm(k,np,npq)


apprx <- function(n, p, R = 1000, k = 6) {
  trueval <- pbinom(k, n, p) # true binomial probability
  prob.zcc <- prob.zncc <- prob.pois <- NULL  
  q<-1-p
  for (i in 1:R) {
    x <- rnorm(n, n * p, sqrt(n * p * q))
    z.cc <- ((k + .5) - mean(x))/sd(x) # with cont. correction
    prob.zcc[i] <- pnorm(z.cc)
    z.ncc <- (k - mean(x))/sd(x) # no cont. correction
    prob.zncc[i] <- pnorm(z.ncc)    
    y <- rpois(n, n * p)
    prob.pois[i] <- length(y[y <= k])/n
  }
 list(prob.zcc = prob.zcc, prob.zncc = prob.zncc, 
       prob.pois = prob.pois, trueval = trueval)
}
R <- 1000
set.seed(10)
out <- apprx(n = 200, p = .03, k = 6, R = 1000)
# windows(6,5)
 plot(1:R, out$prob.pois, type = "l", col = "green", xlab = "Runs", 
      main = expression(paste("Simulated Probabilities: ", 
             n==200, ", ", p==0.03, sep="")),
      ylab = "Probability", ylim = c(.3, .7))
 abline(h = out$trueval, col="red", lty=2)
 lines(1:R, out$prob.zcc, lty = 1, col = "purple")
 lines(1:R, out$prob.zncc, lty = 1, col = "orange")
 legend("bottomleft", c("Poisson", "Normal (with cc)", 
          "Normal (w/o cc)"),
        lty = c(1), col = c("green", "purple", "orange"))


set.seed(10)
out <- apprx(n = 200, p = .03, k = 6, R = 1000)
# windows(6,5)
boxplot(out$prob.pois, boxwex = 0.25, at = 1:1 - .25,
        col = "green",
        main = expression(paste("Approximating Binomial Probability: ", 
                                n==200, ", ", p==0.03, sep="")),
        ylab = "Probablity", 
        ylim = c(out$trueval - 0.2, out$trueval + 0.25))
boxplot(out$prob.zcc, boxwex = 0.25, at = 1:1 + 0, add = T,
         col = "purple")
boxplot(out$prob.zncc, boxwex = 0.25, at = 1:1 + 0.25, add = T,
         col = "orange" )
abline(h = out$trueval, col = "red", lty=2)
legend("topleft", c("Poisson", "Normal (with cc)", "Normal (w/o cc)"), 
           fill = c("green", "purple", "orange"))
```
##  Random variables and their distribution

1. Binomial 

```{r}
dbinom(x=2,size=10,prob=0.4)
pbinom(q=2,size=10,prob=0.4)
qbinom(p=0.4,size=10,prob=0.4)
rbinom(n=5,size=10,prob=0.4)
```

2. The Hypergeometric Distribution

```{r}
dhyper(x=2, m=10, n=30, k=6)
 phyper(q=2, m=10, n=30, k=6)
 qhyper(0.3, m=10, n=30, k=6)
 rhyper(nn=10, m=10, n=30, k=6)
```

3. The Geometric Distribution,let X count the number of failures before the first successes

```{r}
dgeom(4,prob=0.8)
pgeom(4, prob = 0.8)
qgeom(0.4,prob=0.8)
rgeom(10,prob=0.8)
```

4.The Negative Binomial Distribution, let X count the number of failures before r successes

```{r}
dnbinom(x=5,size=3,prob=0.4)   
pnbinom(5,size=3,prob=0.4)
qnbinom(0.5,size=3,prob=0.4)
rnbinom(n=10,size=3,prob=0.4)
```

5. Poisson distributino

```{r}
dpois(x=0,lambda=2.4)
ppois(q=10,lambda=2.4)
qpois(p=0.9,lambda=2.4)
rpois(n=10,lambda=2.4)


 par(mfrow = c(2, 1))
 x <- seq(-0.01, 5, 0.01)
 plot(x, ppois(x, 1), type="s", ylab="F(x)", main="Poisson(1) CDF")
 plot(x, pbinom(x, 100, 0.01),type="s", ylab="F(x)",main="Binomial(100, 0.01) CDF")
```

6. Normal distribution

```{r}
dnorm(0,mean=0,sd=1)
pnorm(0)
qnorm(2.5/100,lower.tail=F)
rnorm(10,mean=1,sd=1.5)
```

some plots

```{r}
x <- seq(-4, 4, length = 401)
plot(x, dnorm(x), type = 'l') # N(0, 1)
# N(1, 1.5^2):
lines(x, dnorm(x, mean = 1, sd = 1.5), lty = 'dashed')
```


```{r}
u <- seq(0, 1, length=401)
plot(u, qnorm(u), 'l')
# lower.tail = FALSE gives q(1-u)
lines(u, qnorm(u, lower.tail = FALSE), lty = 'dashed')
```


```{r}
n<-1000
x<-rnorm(n)
xh<-hist(x, probability = TRUE)
xh
z<-seq(-3,3,0.01)
y <- dnorm(z, mean = 0, sd = 1)
lines(x = z, y = y, col = "blue")
```

7. exponential distribution

```{r}
q = c(.2,.5,.1,.1,.1)
p = 0.5
dexp(x, rate = 1, log = FALSE)
pexp(q, rate = 1, lower.tail = TRUE, log.p = FALSE)
qexp(p, rate = 1, lower.tail = TRUE, log.p = FALSE)
rexp(n, rate = 1)
```

8. Uniform distribution

```{r}
dunif(x, min=0, max=1, log = FALSE)
punif(q, min=0, max=1, lower.tail = TRUE, log.p = FALSE)
qunif(p, min=0, max=1, lower.tail = TRUE, log.p = FALSE)
runif(n, min=0, max=1)
```

9. other distribution


http://zoonek2.free.fr/UNIX/48_R/07.html

## exponential distribution

```{r}
#cumulative distribution function
curve(pexp(x,rate=0.5), xlim=c(0,10), col=1, lwd=3,
      main='Exponential Probability Distribution Function')
curve(pexp(x,rate=1), xlim=c(0,10), col=2, lwd=2, lty=2,
      add=T)
curve(pexp(x,rate=5), xlim=c(0,10), col=3, lwd=2, lty=3,
      add=T)
curve(pexp(x,rate=10), xlim=c(0,10), col=4, lwd=2, lty=4,
      add=T)
legend(par('usr')[2], par('usr')[4], xjust=1,
       c('rate=0.5','rate=1', 'rate=2','rate=10'),
       lwd=2, lty=c(1,2,3,4),
       col=1:4)
```


```{r}
#density
curve(dexp(x,rate=0.5), xlim=c(0,10), col=1, lwd=3,
      main='Exponential Probability Distribution Function')
curve(dexp(x,rate=1), xlim=c(0,10), col=2, lwd=2, lty=2,
      add=T)
curve(dexp(x,rate=5), xlim=c(0,10), col=3, lwd=2, lty=3,
      add=T)
curve(dexp(x,rate=10), xlim=c(0,10), col=4, lwd=2, lty=4,
      add=T)
legend(par('usr')[2], par('usr')[4], xjust=1,
       c('rate=0.5','rate=1', 'rate=2','rate=10'),
       lwd=2, lty=1:4,
       col=1:4)
```


```{r}
###normal
#cumulative distribution function
curve(pnorm(x), xlim=c(-5,5), col='red', lwd=3)
title(main='Cumulative gaussian distribution function')
curve(pnorm(x,1,1), xlim=c(-5,5), col='green', lwd=3,add=T)
curve(pnorm(x,1,2),  xlim=c(-5,5), col='black', lwd=3,add=T)
legend(-par('usr')[2], par('usr')[4], xjust=-0.5,
       c('standard norm', 'normal(1,1)','normal(1,2)'),
       lwd=2, col=c('red','green','black'))
```


```{r}
#density
curve(dnorm(x), xlim=c(-5,5), col='red', lwd=3)
curve(dnorm(x,1,1), add=T, col='green', lty=2, lwd=3)
curve(dnorm(x,1,2), add=T, col='black', lty=3, lwd=3)

legend(par('usr')[2], par('usr')[4], xjust=1,
       c('standard normal', 'normal(1,1)','normal(1,2)'),
       lwd=2, lty=c(1,2,3),
       col=c('red','green','black'))
```


```{r}
###mixture of normal
 
m <- c(-2,0,2)    # Means
p <- c(.3,.4,.3)  # Probabilities
s <- c(1, 1, 1)   # Standard deviations
 
curve( p[2]*dnorm(x, mean=m[2], sd=s[2]),
       col = "green", lwd = 3, 
       xlim = c(-5,5),ylim=c(0,0.23),
       main = "The three gaussian distributions in our mixture",
       xlab = "", ylab = "")
curve( p[1]*dnorm(x, mean=m[1], sd=s[1]),
       col="red", lwd=3, add=TRUE)
curve( p[3]*dnorm(x, mean=m[3], sd=s[3]),
       col="blue", lwd=3, add=TRUE)


curve(p[1]*dnorm(x, mean=m[1], sd=s[1])+p[2]*dnorm(x, mean=m[2], sd=s[2])+p[3]*dnorm(x, mean=m[3], sd=s[3]),col="black", lwd=3, add=TRUE)
```
http://personal.kenyon.edu/hartlaub/MellonProject/Bivariate2.html


```{r eval=FALSE}
### bivariate normal density with matlab


###Plot of mixtures of bivariate normal with R

# install.packages("rgl")
library(rgl)
 
dnorm2d<-function(x,y,mu1,mu2,sigma1,sigma2,rho){
    xoy = ((x-mu1)^2/sigma1^2 - 2*rho * (x-mu1)/sigma1 * (y-mu2)/sigma2 + (y-mu2)^2/sigma2^2)/(2 * (1 - rho^2))
    density = exp(-xoy)/(2 * pi *sigma1*sigma2*sqrt(1 - rho^2))
    density
}



x<-seq(-5,5,by=0.1)
y<-seq(-5,5,by=0.1)

ff1<-function(x,y){0.5*dnorm2d(x,y,0,0,1,1,0)+0.5*dnorm2d(x,y,0,0,1,1,0.5)}

ff2<-function(x,y){0.5*dnorm2d(x,y,0,0,1,1,0.5)+0.5*dnorm2d(x,y,0,0,1,1,-0.5)}


ff3<-function(x,y){0.3*dnorm2d(x,y,0,0,1,1,0)+0.7*dnorm2d(x,y,2.5,2.5,1.75,1.75,0)}

open3d() # This will open a small window where you can plot 3D figures on.

z<-outer(x,y,ff1) 
persp3d(x,y,z,col="green",main="ff1")

open3d()
z<-outer(x,y,ff2)
persp3d(x,y,z,col="green",main="ff2")

open3d()
z<-outer(x,y,ff3)
persp3d(x,y,z,col="green",main="ff3")

 
```

